---
output:
  pdf_document: default
  html_document: default
---
## Analysis of My Course Data

```{r setup, echo=F, message=F, warning=F}
## packages
library(ggplot2)
library(dplyr)
library(xtable)

## setup
theme_set(theme_bw())
knitr::opts_chunk$set(echo = F, message = F, warning = F)

## load data
load("../data/clean-data.rda")
```

Let's start by looking at my performance over time in the two key metrics:

```{r key-metrics}
ggplot(evaluations, aes(x = as.factor(semester_year), y = `The Teaching`)) + 
  geom_jitter(width = 0, alpha = 0.2) +
  stat_summary(color = "red")

ggplot(evaluations, aes(x = as.factor(semester_year), y = `The Course`)) + 
  geom_jitter(width = 0, alpha = 0.2) +
  stat_summary(color = "red")
```

Let's break this down by class:

```{r}
subset(evaluations, course %in% c("211", "301", "311")) %>%
ggplot(aes(x = as.factor(semester_year), y = `The Teaching`)) + 
  geom_jitter(width = 0, alpha = 0.2) +
  stat_summary(color = "red") + 
  facet_wrap(~course, ncol=1)

subset(evaluations, course %in% c("211", "301", "311")) %>%
ggplot(aes(x = as.factor(semester_year), y = `The Course`)) + 
  geom_jitter(width = 0, alpha = 0.2) +
  stat_summary(color = "red") + 
  facet_wrap(~course, ncol=1)
```

These two variables are highly correlated:

```{r}
ggplot(aes(x=`The Course`, y=`The Teaching`), data = evaluations) + 
  geom_jitter(width=.1, height=.1) + 
  stat_smooth(method = "lm")
```

Interestingly, there is not much correlation between the course rating and the reported effort.

```{r}
ggplot(aes(x=`The Course`, y=`On average, how many hours per week did you spend on coursework outside of class?`), data = evaluations) + 
  geom_jitter(width=.1, height=.1) + 
  stat_smooth(method = "loess")
```

It appears that students with lower expected grades work harder in my courses:

```{r}
ggplot(aes(x=`What is your expected grade in this course?`, y=`On average, how many hours per week did you spend on coursework outside of class?`),
       data = subset(evaluations, !is.na(`On average, how many hours per week did you spend on coursework outside of class?`))) +
  stat_boxplot()
```

Let's see what the distribution of grades I give is:

```{r}
xtabs(~ course + Grade, data = subset(enrollments, course %in% c("301", "211", "311")), drop.unused.levels = T)

enrollments %>% 
  filter(course %in% c("301", "211", "311")) %>%
  ggplot(aes(x=Grade)) + geom_bar() + facet_wrap(~course, nrow = 3)

```

Is there any pattern to the average grades in my classes?

```{r}
ggplot(subset(enrollments, course %in% c("301", "211", "311")), aes(x=semester_year, y=GPA)) + 
  stat_summary() +
  facet_wrap(~course)
```


How many enrollments have I taught?

```{r}
dim(enrollments)[1]
```

What is my number of unique students taught?

```{r}
length(unique(enrollments$WesID))
```


```{r}
# potential TAs
# enrollments %>%
#   filter(`Class Year` %in% c(2018, 2019, 2020) & Grade %in% c("A+", "A", "A-", "B+")) %>%
#   select(`E-mail`)
```

